---
title: "Lab9"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

## Read Covid19 Data

First read in the data and convert from wide to long.
```{r, message = F, echo = F}
library(readr)
library(tidyr)
library(dplyr)
library(knitr)
covid_cases <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv") %>% select(-Lat, -Long) %>% rename(state = `Province/State`, country = `Country/Region` )

covid_cases_long <- covid_cases %>% pivot_longer(-c(country,state), names_to = "date", values_to = "count" )
```


### Interactive Table

Allow users to select the date and number of countries to show


```{r covidtable, echo=FALSE}
inputPanel(
  selectInput("date", label = "Date:",
              choices = c( '3/26/20', '3/27/20', '3/28/20','3/29/20','3/30/20'), selected = '3/30/20'))

DT::renderDataTable(
  covid_cases_long %>% group_by(country) %>% filter(date == input$date) %>% group_by(country, date) %>% summarize(`total cases` = sum(count)) %>% arrange(desc(`total cases`)) 
)
```


## Interactive Figure

Filter by state / province and show curves, maybe include colors and/or faceting?

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.


```{r}
library(lubridate)
US_data <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv") %>% select(-c(UID, iso2, iso3, FIPS,Lat,Long_, code3, Country_Region, Combined_Key))

US_long <- US_data %>% pivot_longer(-c(Admin2,Province_State), names_to = "date", values_to = "count" )


```

```{r eruptions, echo=FALSE}
inputPanel(
  
  selectInput("state", label = "State:",
              choices = unique(US_long$Province_State), selected = 'Montana')
)

renderPlot({
  US_long %>% dplyr::filter(Province_State == input$state) %>%  dplyr::mutate(date = mdy(date)) %>% dplyr::group_by(date) %>% dplyr::summarize(count = sum(count)) %>% ggplot2::ggplot(aes(y = count, x = date)) + ggplot2::geom_line() + theme_bw() + ggtitle(paste('Total Diagnosed Covid19 cases in ', input$state ))

})
```


## STATIC MT Plot

```{r, eval = T, echo = F}
library(tidyverse)
library(urbnmapr)

county_data <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv") %>% select(-c(UID, iso2, iso3, Province_State,Lat,Long_, code3, Country_Region, Combined_Key, Admin2))


county_long <- county_data %>% pivot_longer(-c(FIPS), names_to = "date", values_to = "count" ) %>% filter(date == '3/29/20')


df_county <- tibble(FIPS = county.fips$fips)

df_county <- full_join(df_county, county_long, by = "FIPS") %>% filter(!is.na(FIPS), !is.na(count)) 


mt <- urbnmapr::counties %>% filter(state_abbv == 'MT') %>% rename(FIPS = county_fips) %>% mutate(FIPS = as.numeric(FIPS))
mt_map <- left_join(mt, df_county)

mt_map$color <- gray(mt_map$count / max(mt_map$count))


ggplot() +  geom_polygon(data = mt_map, mapping = aes(x = long, y = lat, group = group, fill = count ),
		            color = "gray") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradient(low = "white", high = "red")


```

